---
title: "Lab4"
author: "Akhila Nekkanti"
date: "5/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
#parsed <- rmarkdown::parse_html_notebook("lab4.html")
#str(parsed)
```

# Writing a summary function:
Write a function that takes a given data frame as its first argument, a column to summarize as its second argument, and any set of other columns to summary operations by as the third argument.

## Function behavior:
By default, the function should report:  
* Total number of observations  
* Number of non-missing observations  
* Number of missing observations  
* Mean  
* Standard deviation  
* Minimum  
* Maximum  




arguments- dataset, column name, grouping variable(s) as a list
look into the function summarise_at

?summarise_at
```{r}
library(tidyverse)
View(mtcars)
View(airquality)

#total observations
n <- airquality %>% 
  select(Solar.R) %>% 
  count()

#total non-missing
n_valid <- airquality %>% 
  select(Solar.R) %>% 
  na.omit() %>% 
  count()
n_valid

#total missing
n_missing <- airquality %>% 
  select(Solar.R) %>% 
  summarise(sum(is.na(.)))

### next step is to create a new object that is a list of all the above, use summarise_at to create the list of the functions we need

##actual function is 3 lines of code

function(data, result_var, ...) {
  
}






#function
df_numeric <- function(df) {
    df[ ,map_lgl(df, is.numeric)]
}  

```


### Errors:  
Embed an error if:  
Anything other than a data frame is supplied as the first argument.  
The provided column to summarize is not numeric.  
### Warnings:  
Embed a warning if:  
Any of the grouping variable(s) are numeric (generally you would expect the grouping variable to be a categorical variable)  

### Some guidance:  
If you want to use {dplyr} you’ll need to use non-standard evaluation (NSE), so either something like !!enquo(group_var), where group_var is the formal argument of your function, or {{group_var}}. Also, dplyr::pull (together with NSE) will likely be helpful for checking types for warnings/errors.

Alternatively, you might consider spliting the dataset by the grouping variable, then looping over each split to apply the function(s) supplied.

