---
title: "Joanna_Lab4"
author: "Joanna Wright"
date: "5/20/2020"
output: html_document
---

```{r setup, include=FALSE, error=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE)
library(tidyverse)

```

# Writing a summary function:
Write a function that takes a given data frame as its first argument, a column to summarize as its second argument, and any set of other columns to summary operations by as the third argument. (pass the dots and after, a funs argument that defaults to a list of the functions its going to apply)

## Function behavior:
By default, the function should report:  
* Total number of observations  
* Number of non-missing observations  
* Number of missing observations  
* Mean  
* Standard deviation  
* Minimum  
* Maximum  


```{r}

smry_function <- function(df, outcome, ..., .fun = list(n = function(x) length(x),
                                             n_valid = function(x) sum(!is.na(x)),
                                             n_missing = function(x) sum(is.na(x)),
                                             mean = function(x) mean(x, na.rm = TRUE),
                                             sd = function(x) sd(x, na.rm = TRUE),
                                             min = function(x) min(x, na.rm = TRUE),
                                             max = function(x) max(x, na.rm = TRUE))) {
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

# testing it out: 
# smry_function(gss_cat, tvhours, marital, race, rincome, relig)
# smry_function(gss_cat, tvhours, marital, .fun = list(mean, sd))
# smry_function(airquality, Solar.R, Month, .fun = list(mean))

```

### Errors and warning: 

Embed an error if:  
Anything other than a data frame is supplied as the first argument.  
The provided column to summarize is not numeric. 

Embed a warning if:  
Any of the grouping variable(s) are numeric (generally you would expect the grouping variable to be a categorical variable)  

```{r adding errors and warnings }

smry_function <- function(df, outcome, ..., .fun = list(n = function(x) length(x),
                                             n_valid = function(x) sum(!is.na(x)),
                                             n_missing = function(x) sum(is.na(x)),
                                             mean = function(x) mean(x, na.rm = TRUE),
                                             sd = function(x) sd(x, na.rm = TRUE),
                                             min = function(x) min(x, na.rm = TRUE),
                                             max = function(x) max(x, na.rm = TRUE))) {
  library(dplyr)
            if(!is.data.frame(df)) { 
            stop("Object is not a data frame")
            } 
            if(!is.numeric(pull(df, {{outcome}}))) {
            stop("Outcome is not numeric")
            }
            if(is.numeric(pull(df, ...))) {
            warning("Grouping variables are not categorical")
            }
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

```


Test for df error message: not a data frame 
```{r}
smry_function(smry_function, tvhours, marital)
```


Test for outcome error message: outcome variable not numeric 
```{r}
smry_function(gss_cat, marital, tvhours)
```


Test for warning message: grouping variable not categorical
```{r}
smry_function(gss_cat, tvhours, age) # this one doesn't work yet? 
```


```{r improving error and warning messages}

smry_function2 <- function(df, outcome, ..., .fun = list(n = function(x) length(x),
                                             n_valid = function(x) sum(!is.na(x)),
                                             n_missing = function(x) sum(is.na(x)),
                                             mean = function(x) mean(x, na.rm = TRUE),
                                             sd = function(x) sd(x, na.rm = TRUE),
                                             min = function(x) min(x, na.rm = TRUE),
                                             max = function(x) max(x, na.rm = TRUE))) {
  library(dplyr)
  df_name <- deparse(substitute(df))
  outcome_name <- deparse(substitute(outcome))
  grouping_name <- deparse(substitute(...))
            if(!is.data.frame(df)) { 
            stop(df_name, " is not a data frame")
            } 
            if(!is.numeric(pull(df, {{outcome}}))) {
            stop(outcome_name, " is not numeric")
            }
            if(is.numeric(pull(df, ...))) {
            warning(grouping_name, " Grouping variables are not categorical")
            }
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}


smry_function2(mean, tvhours, marital)

smry_function2(gss_cat, marital, tvhours)

smry_function2(gss_cat, tvhours, age, relig) # not working 

```






