---
title: "Joanna_Lab4"
author: "Joanna Wright"
date: "5/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(dplyr)
```

# Writing a summary function:
Write a function that takes a given data frame as its first argument, a column to summarize as its second argument, and any set of other columns to summary operations by as the third argument. (pass the dots and after, a funs argument that defaults to a list of the functions its going to apply)

## Function behavior:
By default, the function should report:  
* Total number of observations  
* Number of non-missing observations  
* Number of missing observations  
* Mean  
* Standard deviation  
* Minimum  
* Maximum  

However, the function should be flexible so it can take any function(s) as its input to summarize by, provided the function results in a single value (e.g., anything you could feed to dplyr::summarize). You’ll need an additional optional argument (should be the last argument in your function) that lists the function(s) to use in the summary. Hint: You’ll want to have this argument pre-populated with default functions.

```{r}

f1 <- function(x) {
  n <- length(x)
  z <- na.omit(x)
  n_valid <- length(z)
  mean <- mean(z)
  sd <- sd(z)
  min <- min(z)
  max <- max(z)
  return(list(c(n = n, 
                     n_valid =n_valid, 
                     mean=mean,
                     sd=sd,
                     min=min,
                     max=max)))
}


f1(airquality$Solar.R)

tmp <- summarise_at(mtcars, vars(mpg, hp), f1)

```


```{r creating list of functions }

n_fun <- function(data, outcome) {
  data %>%
  select(outcome) %>%
  count()
}
n_fun(airquality, "Solar.R")


n_valid_fun <- function (data, outcome) {
  data %>%
  select(outcome) %>%
  na.omit() %>%
  count()
}
n_valid_fun(airquality, "Solar.R")


n_missing_fun <- function(data, outcome) {
  data %>%
  select(outcome) %>%
  summarise(sum(is.na(.)))
}
n_missing_fun(airquality, "Solar.R")


mean_fun <- function(data, outcome) {
  data %>%
  summarise(mean(outcome, na.rm = TRUE))
}
mean_fun(airquality, airquality$Solar.R)


sd_fun <- function(data, outcome) {
  data %>%
  summarise(sd(outcome, na.rm = TRUE))
}
sd_fun(airquality, airquality$Solar.R)


min_fun <- function(data, outcome) {
  data %>%
  min(outcome, na.rm = TRUE)
}
min_fun(airquality, airquality$Solar.R)


max_fun <- function(data, outcome) {
  data %>%
 max(outcome, na.rm = TRUE)
}
max_fun(airquality, airquality$Solar.R)

```




# create a list
list_tmp <- list(n_fun, n_valid, n_missing, mean_fun, sd_fun, min_fun, max_fun)

# TIPS FROM DA: 
# use summarise_at to pass the list of functions. 
# if that works, then you can get it to work in the body of your function 

# 3rd argument: list of functions. loop through those functions to apply them to the 2nd column




```

### Errors:  
Embed an error if:  
Anything other than a data frame is supplied as the first argument.  
The provided column to summarize is not numeric. 

```{r}

# EXAMPLE ERROR MESSAGE EXAMPLE FROM SLIDES 
both_na <- function(x, y) {
    if(length(x) != length(y)) {
        v_lngths <- paste0("x = ", length(x), ", y = ", length(y))
        stop("Vectors are of different lengths:", v_lngths)
    }
    sum(is.na(x) & is.na(y))
}
both_na(a, c(b, b))

```

### Warnings:  
Embed a warning if:  
Any of the grouping variable(s) are numeric (generally you would expect the grouping variable to be a categorical variable)  

```{r}

#EXAMPLE WARNING MESSAGE FROM SLIDES
#Modify your prior code to so it runs, but returns a warning, if the vectors are recyclable, and otherwise returns a meaningful error message.

#Hint 1: You'll need two conditions

# Hint 2: Check if a number is fractional with %%, which returns the remainder in a division problem. So 8 %% 2 and 8 %% 4 both return zero (because there is no remainder), while and 7 %% 2 returns 1 and 7 %% 4 returns 3.

both_na <- function(x, y) {
    if(length(x) != length(y)) {
        lx <- length(x)
        ly <- length(y)
        v_lngths <- paste0("x = ", lx, ", y = ", ly)
        if(lx %% ly == 0 | ly %% lx == 0) {
            warning("Vectors were recycled (", v_lngths, ")")
        }
        else {
            stop("Vectors are of different lengths and are not recyclable:",
                 v_lngths)    
        }
    }
    sum(is.na(x) & is.na(y))
}

```


### Some guidance:  
If you want to use {dplyr} you’ll need to use non-standard evaluation (NSE), so either something like !!enquo(group_var), where group_var is the formal argument of your function, or {{group_var}}. Also, dplyr::pull (together with NSE) will likely be helpful for checking types for warnings/errors.

Alternatively, you might consider spliting the dataset by the grouping variable, then looping over each split to apply the function(s) supplied.


```{r}
# other random notes/examples

# ways to specify return:

add_two.1 <- function(x) {
    result <- x + 2
    result
}

add_two.2 <- function(x) {
    x + 2
}

add_two.3 <- function(x) {
    result <- x + 2
    return(result)
}

# example function
lm_model <- function(data, outcome) {
  lm(data[[outcome]] ~ data[["total_cases"]])
}
# testing it on one category, "parks": 
lm_model(d, "parks") %>% 
  return()


# incomplete template? 
function_name <- function(arg1, arg2) {
  data %>%
    summarize(min = min(col),
              max = max(col),
              mean = mean(col),
              sd = sd(col),
    mutate_if(is.numeric, round, 2)
    )
}



# 3rd argument: list of functions. loop through those functions to apply them to the 2nd column


