---
title: "Rowley_Lab4"
author: "Brock Rowley"
date: "5/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
#parsed <- rmarkdown::parse_html_notebook("lab4.html")
#str(parsed)
```

# Writing a summary function:
Write a function that takes a given data frame as its first argument, a column to summarize as its second argument, and any set of other columns to summary operations by as the third argument.

## Function behavior:
By default, the function should report:  
* Total number of observations  
* Number of non-missing observations  
* Number of missing observations  
* Mean  
* Standard deviation  
* Minimum  
* Maximum  
However, the function should be flexible so it can take any function(s) as its input to summarize by, provided the function results in a single value (e.g., anything you could feed to dplyr::summarize). You’ll need an additional optional argument (should be the last argument in your function) that lists the function(s) to use in the summary. Hint: You’ll want to have this argument pre-populated with default functions.

### Errors:  
Embed an error if:  
Anything other than a data frame is supplied as the first argument.  
The provided column to summarize is not numeric.  
### Warnings:  
Embed a warning if:  
Any of the grouping variable(s) are numeric (generally you would expect the grouping variable to be a categorical variable)  
```{r sum_fun}
# smry <- list(n = length, 
#              n_miss = function(x) sum(is.na(x)),
#              n_valid = function(x) sum(!is.na(x)),
#              mean = mean, 
#              sd = sd)

# summarize_col <- function(column) {
#   map_df(smry, ~.x(column))
#}

# summarize_df <- function(df) {
#   map_df(df, summarize_col, .id = "column")
# }

summarize_cols <- function(data, ...) {
    data %>%
        select(...) %>%
        pivot_longer(everything(), names_to = "var", values_to = "val") %>%
        group_by(var) %>%
        summarize(mean = mean(val, na.rm = TRUE),
                  sd = sd(val, na.rm = TRUE),
                  min = min(val, na.rm = TRUE),
                  max = max(val, na.rm = TRUE))
}

my_mean <- function(x) {
	if(any(is.na(x))) {
		tot_na <- sum(is.na(x))
		warning(tot_na, " observations removed due to missingness")
		x <- na.omit(x)
	}
	sum(x) / length(x)
}

df_summary <- function(df, as_tibble = TRUE) {
	df_num <- df_numeric(df)
	means <- purrr::map_dbl(df_num, my_mean)
	sds <- purrr::map_dbl(df_num, sd, na.rm = TRUE)

	if(as_tibble) {
		tbl <- tibble::tibble(variable = names(df_num),
		                      mean = means,
		                      sd = sds)
	} else {
		tbl <- data.frame(variable = names(df_num),
		                  mean = means,
		                  sd = sds)
	}
	tbl
}
```

### Some guidance:  
If you want to use {dplyr} you’ll need to use non-standard evaluation (NSE), so either something like !!enquo(group_var), where group_var is the formal argument of your function, or {{group_var}}. Also, dplyr::pull (together with NSE) will likely be helpful for checking types for warnings/errors.

Alternatively, you might consider spliting the dataset by the grouping variable, then looping over each split to apply the function(s) supplied.

### Sample output:
Below are some examples of how the function should work. Note that I have purposefully given my function a poor name, but you should think carefully about the name of your function.

Examples with {dplyr}:
<p><strong>Examples with {dplyr}:</strong></p>
<pre class="r"><code>f(iris, Sepal.Width, Species)</code></pre>
<pre><code>## # A tibble: 3 x 8
##   Species        n n_valid n_miss  mean        sd      min   max
##   &lt;fct&gt;      &lt;int&gt;   &lt;int&gt;  &lt;int&gt; &lt;dbl&gt;     &lt;dbl&gt;    &lt;dbl&gt; &lt;dbl&gt;
## 1 setosa        50      50      0 3.428 0.3790644 2.300000   4.4
## 2 versicolor    50      50      0 2.77  0.3137983 2          3.4
## 3 virginica     50      50      0 2.974 0.3224966 2.2        3.8</code></pre>
<pre class="r"><code>f(airquality, Ozone, Month)</code></pre>
<pre><code>## Warning in f(airquality, Ozone, Month): This is a warning about a numeric
## grouping variable. Make your message more informative.</code></pre>
<pre><code>## # A tibble: 5 x 8
##   Month     n n_valid n_miss     mean       sd   min   max
##   &lt;int&gt; &lt;int&gt;   &lt;int&gt;  &lt;int&gt;    &lt;dbl&gt;    &lt;dbl&gt; &lt;int&gt; &lt;int&gt;
## 1     5    31      26      5 23.61538 22.22445     1   115
## 2     6    30       9     21 29.44444 18.20790    12    71
## 3     7    31      26      5 59.11538 31.63584     7   135
## 4     8    31      26      5 59.96154 39.68121     9   168
## 5     9    30      29      1 31.44828 24.14182     7    96</code></pre>
<pre class="r"><code>airquality %&gt;%
    mutate(Month = as.factor(Month)) %&gt;%
    f(Ozone, Month)</code></pre>
<pre><code>## # A tibble: 5 x 8
##   Month     n n_valid n_miss     mean       sd   min   max
##   &lt;fct&gt; &lt;int&gt;   &lt;int&gt;  &lt;int&gt;    &lt;dbl&gt;    &lt;dbl&gt; &lt;int&gt; &lt;int&gt;
## 1 5        31      26      5 23.61538 22.22445     1   115
## 2 6        30       9     21 29.44444 18.20790    12    71
## 3 7        31      26      5 59.11538 31.63584     7   135
## 4 8        31      26      5 59.96154 39.68121     9   168
## 5 9        30      29      1 31.44828 24.14182     7    96</code></pre>
<pre class="r"><code>f(iris$Sepal.Width, iris$Species)</code></pre>
<pre><code>## Error in f(iris$Sepal.Width, iris$Species): This is a non-df error. Make your message more informative</code></pre>
<pre class="r"><code>mtcars2 &lt;- mtcars %&gt;%
  mutate_at(.vars = vars(cyl, vs, am, gear, carb),
            .funs = as.factor)

f(mtcars2, mpg, cyl, vs, am)</code></pre>
<pre><code>## # A tibble: 7 x 10
## # Groups:   cyl, vs [5]
##   cyl   vs    am        n n_valid n_miss     mean         sd   min   max
##   &lt;fct&gt; &lt;fct&gt; &lt;fct&gt; &lt;int&gt;   &lt;int&gt;  &lt;int&gt;    &lt;dbl&gt;      &lt;dbl&gt; &lt;dbl&gt; &lt;dbl&gt;
## 1 4     0     1         1       1      0 26       NA          26    26  
## 2 4     1     0         3       3      0 22.9      1.452584   21.5  24.4
## 3 4     1     1         7       7      0 28.37143  4.757701   21.4  33.9
## 4 6     0     1         3       3      0 20.56667  0.7505553  19.7  21  
## 5 6     1     0         4       4      0 19.125    1.631717   17.8  21.4
## 6 8     0     0        12      12      0 15.05     2.774396   10.4  19.2
## 7 8     0     1         2       2      0 15.4      0.5656854  15    15.8</code></pre>
<pre class="r"><code>f(mtcars2, mpg, cyl, vs, am, .funs = median)</code></pre>
<pre><code>## # A tibble: 7 x 4
## # Groups:   cyl, vs [5]
##   cyl   vs    am       mpg
##   &lt;fct&gt; &lt;fct&gt; &lt;fct&gt;  &lt;dbl&gt;
## 1 4     0     1     26    
## 2 4     1     0     22.8  
## 3 4     1     1     30.4  
## 4 6     0     1     21    
## 5 6     1     0     18.650
## 6 8     0     0     15.2  
## 7 8     0     1     15.4</code></pre>
<pre class="r"><code>f(mtcars2, mpg, cyl, vs, am, .funs = list(~median(.), 
                                            ~quantile(., 0.2)))</code></pre>
<pre><code>## # A tibble: 7 x 5
## # Groups:   cyl, vs [5]
##   cyl   vs    am    median quantile
##   &lt;fct&gt; &lt;fct&gt; &lt;fct&gt;  &lt;dbl&gt;    &lt;dbl&gt;
## 1 4     0     1     26        26   
## 2 4     1     0     22.8      22.02
## 3 4     1     1     30.4      23.7 
## 4 6     0     1     21        20.22
## 5 6     1     0     18.650    17.98
## 6 8     0     0     15.2      13.5 
## 7 8     0     1     15.4      15.16</code></pre>
<pre class="r"><code>f(mtcars2, mpg, cyl, vs, am, .funs = list(med = ~median(.), 
                                          q20 = ~quantile(., 0.2)))</code></pre>
<pre><code>## # A tibble: 7 x 5
## # Groups:   cyl, vs [5]
##   cyl   vs    am       med   q20
##   &lt;fct&gt; &lt;fct&gt; &lt;fct&gt;  &lt;dbl&gt; &lt;dbl&gt;
## 1 4     0     1     26     26   
## 2 4     1     0     22.8   22.02
## 3 4     1     1     30.4   23.7 
## 4 6     0     1     21     20.22
## 5 6     1     0     18.650 17.98
## 6 8     0     0     15.2   13.5 
## 7 8     0     1     15.4   15.16</code></pre>

<p><strong>Examples without {dplyr}</strong></p>
<p>Note the syntax will look slightly different (because you’re not using NSE) but it all should still work basically the same.</p>
<pre class="r"><code>f2(mtcars2, &quot;mpg&quot;, &quot;cyl&quot;, &quot;vs&quot;, &quot;am&quot;)</code></pre>
<pre><code>##   cyl vs am  n n_valid n_miss     mean        sd  min  max
## 1   8  0  0 12      12      0 15.05000 2.7743959 10.4 19.2
## 2   4  1  0  3       3      0 22.90000 1.4525839 21.5 24.4
## 3   6  1  0  4       4      0 19.12500 1.6317169 17.8 21.4
## 4   4  0  1  1       1      0 26.00000        NA 26.0 26.0
## 5   6  0  1  3       3      0 20.56667 0.7505553 19.7 21.0
## 6   8  0  1  2       2      0 15.40000 0.5656854 15.0 15.8
## 7   4  1  1  7       7      0 28.37143 4.7577005 21.4 33.9</code></pre>
<pre class="r"><code>f2(iris, &quot;Sepal.Width&quot;, &quot;Species&quot;, .funs = list(med = median,
                                                q20 = function(x) quantile(x, 0.2)))</code></pre>
<pre><code>##      Species med q20
## 1     setosa 3.4 3.1
## 2 versicolor 2.8 2.5
## 3  virginica 3.0 2.7</code></pre>