---
title: "Rowley_Lab4"
author: "Brock Rowley"
date: "5/19/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Writing a summary function:
Write a function that takes a given data frame as its first argument, a column to summarize as its second argument, and any set of other columns to summary operations by as the third argument.

## Function behavior:
By default, the function should report:  
* Total number of observations  
* Number of non-missing observations  
* Number of missing observations  
* Mean  
* Standard deviation  
* Minimum  
* Maximum  
However, the function should be flexible so it can take any function(s) as its input to summarize by, provided the function results in a single value (e.g., anything you could feed to dplyr::summarize). You’ll need an additional optional argument (should be the last argument in your function) that lists the function(s) to use in the summary. Hint: You’ll want to have this argument pre-populated with default functions.

```{r summary_function}
smry_function <- function(
  df, outcome, ..., .fun = list(
    n = function(x) length(x),
    n_valid = function(x) sum(!is.na(x)),
    n_missing = function(x) sum(is.na(x)),
    mean = function(x) mean(x, na.rm = TRUE),
    sd = function(x) sd(x, na.rm = TRUE),
    min = function(x) min(x, na.rm = TRUE),
    max = function(x) max(x, na.rm = TRUE))) {
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

#testing it out on 2 different datasets
smry_function(iris, Sepal.Width, Species)
smry_function(gss_cat, tvhours, relig, race)


listof_fun <- list(
  n = function(x) length(x),
  n_valid = function(x) sum(!is.na(x)),
  n_missing = function(x) sum(is.na(x)),
  mean = function(x) mean(x, na.rm = TRUE),
  sd = function(x) sd(x, na.rm = TRUE),
  min = function(x) min(x, na.rm = TRUE),
  max = function(x) max(x, na.rm = TRUE)
)

smry_function2 <- function(df, outcome, ..., .fun = listof_fun
) {
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

#testing it out on 2 different datasets
smry_function2(iris, Sepal.Width, Species)
smry_function2(airquality, Wind, Ozone, Month)
```

### Errors:  
Embed an error if:  
Anything other than a data frame is supplied as the first argument.  
The provided column to summarize is not numeric.  

### Warnings:  
Embed a warning if:  
Any of the grouping variable(s) are numeric (generally you would expect the grouping variable to be a categorical variable)  

## Generic Errors and Warnings:
Our warning message for whether grouping variables were categorical does NOT work when you supply more than one grouping variable, see test below.

```{r error_warn, error = TRUE, warnings = TRUE}
check_numeric <- function(df, ...){
  colums <- select(df, ...)
  any(map_lgl(colums, is.numeric))
}

smry_function <- function(
  df, outcome, ..., .fun = list(
    n = function(x) length(x),
    n_valid = function(x) sum(!is.na(x)),
    n_missing = function(x) sum(is.na(x)),
    mean = function(x) mean(x, na.rm = TRUE),
    sd = function(x) sd(x, na.rm = TRUE),
    min = function(x) min(x, na.rm = TRUE),
    max = function(x) max(x, na.rm = TRUE))) {
            if (!is.data.frame(df)) { 
            stop("Object is not a data frame")
            } 
            if (!is.numeric(pull(df, {{outcome}}))) {
            stop("Outcome is not numeric")
            }
            if (check_numeric(df, ...)) {
            warning("Grouping variables are not categorical")
            }
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

#test for df 
smry_function(smry_function, tvhours, marital)

#test for outcome numeric
smry_function(gss_cat, marital, tvhours)

#test for grouping variable categorical 
smry_function(gss_cat, tvhours, age)

# test for grouping variable categorical (multiple grouping variables) 
smry_function(gss_cat, tvhours, age, year)
smry_function(airquality, Wind, Ozone, Month)

smry_function2(gss_cat, tvhours, age, year)
smry_function2(airquality, Wind, Ozone, Month)
```

## Errors and Warnings Customized:
Our attempt at customizing errors and warnings. We were able to get the data frame and is numeric error to print the correct names for df and outcome. Grouping variable prints correctly for 1 variable, but was NOT able to get the grouping variable printed correctly for multiple variables. Similar to above, if supplying more than 1 grouping variable, then the fuction does NOT work with warnings. 

```{r errors_warnings, error = TRUE, warnings = TRUE}

smry_function_custom <- function(
  df, outcome, ..., .fun = list(
    n = function(x) length(x),
    n_valid = function(x) sum(!is.na(x)),
    n_missing = function(x) sum(is.na(x)),
    mean = function(x) mean(x, na.rm = TRUE),
    sd = function(x) sd(x, na.rm = TRUE),
    min = function(x) min(x, na.rm = TRUE),
    max = function(x) max(x, na.rm = TRUE))) {
  df_name <- deparse(substitute(df))
  outcome_name <- deparse(substitute(outcome))
  grouping_name <- sapply(substitute(...()), deparse)
            if (!is.data.frame(df)) { 
            stop(df_name, " is not a data frame")
            } 
            if (!is.numeric(pull(df, {{outcome}}))) {
            stop(outcome_name, " is not numeric")
            }
            if (check_numeric(df, ...)) {
            warning(grouping_name, " grouping variables are not categorical")
            }
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

#test for df 
smry_function_custom(smry_function, tvhours, marital)

#test for outcome numeric
smry_function_custom(gss_cat, marital, tvhours)

#test for grouping variable categorical, 1 variable, works and produces correct warning! 
smry_function_custom(gss_cat, tvhours, year)

#test for grouping variable categorical, multiple variables (does NOT work) 
smry_function_custom(gss_cat, tvhours, year, age)
```