---
title: "Fun Programming Lab 4"
author: "Stephanie Gluck"
date: "5/25/2020"
output: html_document
---


```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Writing a summary function:
Write a function that takes a given data frame as its first argument, a column to summarize as its second argument, and any set of other columns to summary operations by as the third argument.

## Function behavior:
By default, the function should report:  
* Total number of observations  
* Number of non-missing observations  
* Number of missing observations  
* Mean  
* Standard deviation  
* Minimum  
* Maximum  

However, the function should be flexible so it can take any function(s) as its input to summarize by, provided the function results in a single value (e.g., anything you could feed to dplyr::summarize). You’ll need an additional optional argument (should be the last argument in your function) that lists the function(s) to use in the summary. Hint: You’ll want to have this argument pre-populated with default functions.

## Our custom function 

**Stephanie comment:** Our group was not able to get the function to run only specifcied functions such as only returning n_valid and n_missing. 

```{r summary_function}

smry_function <- function(df, outcome, ..., .fun = list(n = function(x) length(x),
                                             n_valid = function(x) sum(!is.na(x)),
                                             n_missing = function(x) sum(is.na(x)),
                                             mean = function(x) mean(x, na.rm = TRUE),
                                             sd = function(x) sd(x, na.rm = TRUE),
                                             min = function(x) min(x, na.rm = TRUE),
                                             max = function(x) max(x, na.rm = TRUE))) {
  library(dplyr)
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

#testing it out on 2 different datasets
smry_function(iris, Sepal.Width, Species)
smry_function(gss_cat, tvhours, relig, race)


#not standalone option, if you create a list of function to later call into the summary function. 

listof_fun <- list(
  n = function(x) length(x),
  n_valid = function(x) sum(!is.na(x)),
  n_missing = function(x) sum(is.na(x)),
  mean = function(x) mean(x, na.rm = TRUE),
  sd = function(x) sd(x, na.rm = TRUE),
  min = function(x) min(x, na.rm = TRUE),
  max = function(x) max(x, na.rm = TRUE)
)

smry_function2 <- function(df, outcome, ..., .fun = listof_fun
) {
  library(dplyr)
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

#testing it out on 2 different datasets
smry_function2(iris, Sepal.Width, Species)
smry_function2(airquality, Wind, Ozone, Month)


```


### Errors:  
Embed an error if:  
Anything other than a data frame is supplied as the first argument.  
The provided column to summarize is not numeric.  


### Warnings:  
Embed a warning if:  
Any of the grouping variable(s) are numeric (generally you would expect the grouping variable to be a categorical variable)  


## Generic errors and warnings

**Stephanie comment:** Our warning message for whether grouping variables were categorical does NOT work when you supply more than one grouping variable, see test below.


```{r errors_warning, error = TRUE, warnings = TRUE}

#Our first take on errors and warnings where we supplied generic error and warning messages. 

smry_function <- function(df, outcome, ..., .fun = list(n = function(x) length(x),
                                             n_valid = function(x) sum(!is.na(x)),
                                             n_missing = function(x) sum(is.na(x)),
                                             mean = function(x) mean(x, na.rm = TRUE),
                                             sd = function(x) sd(x, na.rm = TRUE),
                                             min = function(x) min(x, na.rm = TRUE),
                                             max = function(x) max(x, na.rm = TRUE))) {
  library(dplyr)
            if(!is.data.frame(df)) { 
            stop("Object is not a data frame")
            } 
            if(!is.numeric(pull(df, {{outcome}}))) {
            stop("Outcome is not numeric")
            }
            if(is.numeric(pull(df, ...))) {
            warning("Grouping variables are not categorical")
            }
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

#test for df 
smry_function(smry_function, tvhours, marital)

#test for outcome numeric
smry_function(gss_cat, marital, tvhours)

#test for grouping variable categorical 
smry_function(gss_cat, tvhours, age)

#test for grouping variable categorical (multiple grouping variables) 
#DOES NOT WORK! 
smry_function(gss_cat, tvhours, age, year)
smry_function(airquality, Wind, Ozone, Month)

#but original funciton does work for multiple variable, but function with error and warning does NOT work 
smry_function2(gss_cat, tvhours, age, year)
smry_function2(airquality, Wind, Ozone, Month)


```


## Customized errors and warnings

**Stephanie comment:** Our attempt at customizing errors and warnings. We were able to get the data frame and is numeric error to print the correct names for df and outcome. Grouping variable prints correctly for 1 variable, but was NOT able to get the grouping variable printed correctly for multiple variables. Similar to above, if supplying more than 1 grouping variable, then the fuction does NOT work with warnings. 

```{r errors_warnings_custom, error = TRUE, warnings = TRUE}

smry_function_custom <- function(df, outcome, ..., .fun = list(n = function(x) length(x),
                                             n_valid = function(x) sum(!is.na(x)),
                                             n_missing = function(x) sum(is.na(x)),
                                             mean = function(x) mean(x, na.rm = TRUE),
                                             sd = function(x) sd(x, na.rm = TRUE),
                                             min = function(x) min(x, na.rm = TRUE),
                                             max = function(x) max(x, na.rm = TRUE))) {
  library(dplyr)
  df_name <- deparse(substitute(df))
  outcome_name <- deparse(substitute(outcome))
  #grouping_name <- deparse(substitute(...))
  grouping_name <- sapply(substitute(...()), deparse)
            if(!is.data.frame(df)) { 
            stop(df_name, " is not a data frame")
            } 
            if(!is.numeric(pull(df, {{outcome}}))) {
            stop(outcome_name, " is not numeric")
            }
            if(is.numeric(pull(df, ...))) {
            warning(grouping_name, " grouping variables are not categorical")
            }
  df %>% 
    group_by(...) %>% 
    summarise_at(vars({{outcome}}), .funs = .fun)
}

#test for df 
smry_function_custom(smry_function, tvhours, marital)

#test for outcome numeric
smry_function_custom(gss_cat, marital, tvhours)

#test for grouping variable categorical, 1 variable, works and produces correct warning! 
smry_function_custom(gss_cat, tvhours, year)

#test for grouping variable categorical, multiple variables (does NOT work) 
smry_function_custom(gss_cat, tvhours, year, age)


```


### Some guidance:  
If you want to use {dplyr} you’ll need to use non-standard evaluation (NSE), so either something like !!enquo(group_var), where group_var is the formal argument of your function, or {{group_var}}. Also, dplyr::pull (together with NSE) will likely be helpful for checking types for warnings/errors.

Alternatively, you might consider spliting the dataset by the grouping variable, then looping over each split to apply the function(s) supplied.


### Extra Notes 

**Stephanie comment: Please ignore for grading of Lab 4.** These were codes that helped our group get to the final codes for lab 4. I am keeping them here as notes to myself for record keeping purpose to help with understanding our thought process for the lab. 


```{r notes, eval = FALSE}

#function(data, results_var, ...){}

# Start with how we'd obtain summaries usind dplyr or base R for one dataset and column of interest 

n <- airquality %>%
  select(Solar.R) %>%
  count()

n

n_valid <- airquality %>%
  select(Solar.R) %>%
  na.omit() %>%
  count()

n_valid

n_missing <- airquality %>%
  select(Solar.R) %>%
  summarise(sum(is.na(.)))

n_missing

mean1 <- airquality %>%
  na.omit() %>%
  summarise(mean(Solar.R))

sd <- airquality %>%
  na.omit() %>%
  summarise(sd(Solar.R))

sd

min <- airquality %>%
  na.omit() %>%
  summarise(min = min(Solar.R))

min

max <- airquality %>%
  na.omit() %>%
  summarise(max = max(Solar.R))

max

# generalize the above to be function for any dataset

# can add curly bracket for NSE of the outcome if don't want to have to add quote mark to your outcome variable. 
n_fun <- function(data, outcome) {
  data %>%
  select({{outcome}}) %>%
  count()
}

n_fun(airquality, Solar.R)

# if no curly bracket around outcome, then have to use quote marks on your outcome variable. 
# n_valid_fun <- function (data, outcome) {
#   data %>%
#   select(outcome) %>%
#   na.omit() %>%
#   count()
# }

n_valid_fun <- function (data, ...) {
  data %>%
  select(...) %>%
  na.omit() %>%
  count()
}

n_valid_fun(airquality, Solar.R)

n_missing_fun <- function(data, ...) {
  data %>%
  select(...) %>%
  summarise(sum(is.na(.)))
}

n_missing_fun(airquality, Solar.R)

mean_fun <- function(data, outcome) { 
  data %>%
  na.omit() %>%
  summarise(mean({{outcome}}))
}

mean_fun(airquality, Solar.R)


sd_fun <- function(data, outcome) {
  data %>%
  na.omit() %>%
  summarise(sd({{outcome}}))
}

sd_fun(airquality, Solar.R)


min_fun <- function(data, outcome) {
  data %>%
  na.omit() %>%
  summarise(min = min({{outcome}}))
}

min_fun(airquality, Solar.R)

max_fun <- function(data, outcome) {
  data %>%
  na.omit() %>%
  summarise(max = max({{outcome}}))
}

max_fun(airquality, Solar.R)

# using summarise_at 
# summarise the list of columns of dataframe
library(dplyr)
summarise_at(mtcars, vars(mpg, hp), list(mean = mean, median = median))


# TIPS FROM DA: 
# use summarise_at to pass the list of functions. 
# if that works, then you can get it to work in the body of your function 
# 3rd argument: list of functions. loop through those functions to apply them to the 2nd column

# ... is group by 
# want to group by all these variables and summarize them 
# apply function from your list 
# df, group by , summarize (but summarize is dynamic and apply a bunch of function from a list)

```

